---
title: "Overview of gTrack parameters"
output: html_document
---

```{r include = FALSE, message = FALSE, warning = FALSE}
devtools::load_all("~/git/gTrack")
library(gTrack)
library(gUtils)
knitr::opts_chunk$set(fig.height=10)
```

This is a (non-comprehensive) overview of `gTrack` parameters that are commonly modified from their default values.

# Axes

## Track names (name)

The `name` specifies the track label that is shown to the right of a subplot. Arguments must be character vectors. The default is an empty string, which creates an unlabeled plot. Below is an example of a scatter plot with and without a `name` argument.

```{r include = TRUE, message = FALSE, warning = FALSE}
coverage.gr = readRDS(system.file("extdata", "ovcar.subgraph.coverage.rds", package = "gTrack"))

## read depth gTrack without name argument
coverage.gt = gTrack(coverage.gr, y.field = "cn", circles = TRUE, lwd.border = 0.2, y0 = 0, y1 = 12)

## read depth gTrack with name argument
coverage.name.gt = gTrack(coverage.gr, name = "coverage", y.field = "cn", circles = TRUE, lwd.border = 0.2, y0 = 0, y1 = 12)

fp = parse.gr("1:6043576-7172800")
plot(c(coverage.gt, coverage.name.gt), fp + 1e5)
```

## Quantile-based Y-axis cutoffs (y.quantile)

By default, the minimum and maximum y-axis values are determined by the values of the 1st and 99th percentiles of the data to be plotted. This percentile can be changed by adjusting `y.quantile` in order to plot a wider or narrower range of points. The example below compares two different values of `y.quantile`.

```{r include = TRUE, message = FALSE, warning = FALSE}
coverage.gr = readRDS(system.file("extdata", "ovcar.subgraph.coverage.rds", package = "gTrack"))

## plot points between 1st and 99th percentile
coverage.1.gt = gTrack(coverage.gr, y.field = "cn", circles = TRUE, lwd.border = 0.2, y.quantile = 0.01)

## plot points between 20th and 80th percentile
coverage.20.gt = gTrack(coverage.gr, y.field = "cn", circles = TRUE, lwd.border = 0.2, y.quantile = 0.2)

fp = parse.gr("1:6043576-7172800")
plot(c(coverage.1.gt, coverage.20.gt), fp + 1e5)
```

## Specifying Y-axis cutoffs (y0 and y1)

The minimum and maximum y-axis values can also be specified directly with `y0` and `y1` respectively, which must be numeric values. Below is an example of the same data with two different settings for `y0` and `y1`.

```{r include = TRUE, message = FALSE, warning = FALSE}
coverage.gr = readRDS(system.file("extdata", "ovcar.subgraph.coverage.rds", package = "gTrack"))

## plot points between 0 and 12
coverage.wide.gt = gTrack(coverage.gr, y.field = "cn", circles = TRUE, lwd.border = 0.2, y0 = 0, y1 = 12)

## plot points between 5 and 10
coverage.narrow.gt = gTrack(coverage.gr, y.field = "cn", circles = TRUE, lwd.border = 0.2, y0 = 5, y1 = 10)

fp = parse.gr("1:6043576-7172800")
plot(c(coverage.wide.gt, coverage.narrow.gt), fp + 1e5)
```

## Unbounded Y-axes (y.cap)

## Y-axis labels (yaxis)

# Point density and spacing

## Stacking adjacent ranges (stack.gap)

## Number of plotted ranges (max.ranges)

# Data labels

# Point and line style

## Scatterplot point size (lwd.border)

## Width of plotted ranges (ywid)

# Colors

## Specifying data color directly (col)

## Coloring by secondary label (gr.colorfield)

## Custom map from label to color (colormaps)

## Background color (bg.col)

## Color thresholds (cmap.max and cmap.min)

# Legends
